Rotation Period Evolution in Low-Mass Binary Stars: The Impact of Tidal Torques and Magnetic Braking

### 2019-02-04 Paper/sync.tex
Updates from ShareLaTeX

Tidal torques pose a fundamental problem for inferring ages of stars via gyrochronlogy. Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$ and primary star masses, decoupling P$_{rot}$ from age. For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages $\lsim 2$ Gyr, e.g. the right panel of Fig.~\ref{fig:protDist} or Fig.~\ref{fig:protDistReiners}.  If that system is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail. This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal.

>>>>

Tidal torques pose a fundamental problem for inferring ages of stars via gyrochronlogy. Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$ and primary star masses, decoupling P$_{rot}$ from age. For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages $\lsim 2$ Gyr. We highlight this point in Fig.~\ref{fig:ageHist} by plotting a histogram of system ages from Fig.~\ref{fig:protDist} for single or primary stars in binaries with P$_{rot} < 20$ d.  If a rapidly-rotating star is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail as its age could be at least several Gyrs. This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), e.g. Fig.~\ref{fig:protAgeHist}, but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal, e.g. Fig.~\ref{fig:protDist} and Fig.~\ref{fig:protDistReiners}.

\begin{figure}[h]
	\includegraphics[width=0.45\textwidth]{../Plots/protAgeHist.pdf}
   \caption{Histogram of system ages for stars with P$_{rot} < 20$ d evolved using the \citet{Matt2015} magnetic braking model.}%
    \label{fig:protAgeHist}%
\end{figure}
